### 2017-03-01 doc/Thesis/data/dynamics.tex
Thesis: adding conclusion
more comments pushed. almost there

Reactor dynamics is the study of the time-dependent behavior of reactors as an entire system. This study includes the physical nature of neutrons and their feedback with power generation.  This feedback includes coupling with other physical properties such as temperature, fluids, material dynamics, etc.  This section describes several dynamics examples, including the LRA benchmark and two TREAT experiments.  These examples are of increased complexity from the kinetics examples of \sct{sect:kin}.  This section also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.

\section{LRA Benchmark}

The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.  It is a super prompt-critical transient.  To have better understanding on the cross sections given later, we present the equations here:
\begin{subequations}
\begin{align}
-\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
P &= \kappa S_f,
\end{align}
\end{subequations}
where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
\begin{align}
-\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
-\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1.
\end{align}
The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.  The initial flux distribution shall be normalized such that the averaged power density
\begin{align}
\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
\end{align}
where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\

@@

\subsection{LRA Multiphysics Time Scale Results}

@@

\subsection{LRA with Time Adaptation}

@@

\section{TREAT Transient-15 Problem}

Transient 15 is a test case based on the TREAT core. The preliminary purpose of this model was to match an early test of TREAT; due to lack of experimental data and procedures, model validation was impossible \cite{mammoth}. Therefore, ultimate purpose of this simulation in Rattlesnake is to test the model's fidelity with the thermal feedback of TREAT, but it is not meant to exactly match any previous experiments.  Nevertheless, the goal of the following simulations is to test IQS and its time scale based treatment of temperature with a more complex model. The model involves a 159-element "small core" configuration of TREAT, shown in \fig{fig:Tran15_config} \cite{Tran15}. \fig{fig:Tran15_mesh} shows the meshing techniques of the model \cite{mammoth}. To ease computation, the fully homogenized version (without air gaps) was used  for the following simulations. Transient 15 involves an 11-energy group diffusion approximation and is discretized into $355,712$ hexahedral continuous finite elements totaling $4,109,523$ degrees of freedom.  The three-second transient involves a linear ramp decrease in the absorption cross section throughout the control rod region. \fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\linewidth]{\FiguresDir/Tran15_config.png}
\caption{Transient 15 159-element small core configuration}
\label{fig:Tran15_config}
\end{figure}

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/Tran15_mesh_fine.png}
\caption{Explicit meshing of air channels}
\label{fig:Tran15_mesh_fine}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/Tran15_mesh_homo.png}
\caption{Fully homogenized fuel elements}
\label{fig:Tran15_mesh_homo}
\end{subfigure}
\caption{Top quarter view of Transient 15 mesh}
\label{fig:Tran15_mesh}
\end{figure}

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.6\linewidth]{\FiguresDir/Tran15_core2.png}
\caption{Transient 15 core power profile at peak power}
\label{fig:Tran15}
\end{figure}

\subsection{Transient-15 Temperature Feedback}

@@

\be
c_p = -5.8219\text{e-}10T^3 - 4.3694\text{e-}7T^2 + 2.8369\text{e-}3T -1.009\text{e-}2 \ (J/g/K)
\label{eq:cp}
\ee

\subsection{Transient-15 Multiphysics Time Scale Results}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{\FiguresDir/Tran15_profile.png}
\caption{Transient 15 total power and average temperature profile during transient}
\label{fig:Tran15_profile}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\linewidth]{\FiguresDir/time_constant_tran15.png}
\caption{Dynamical time scale for the Transient 15 example}
\label{fig:tran15tc}
\end{figure}

\subsection{Transient-15 with Time Adaptation}

@@

\section{TREAT M8-CAL Problem}

>>>>

Reactor dynamics is the study of the time-dependent behavior of reactors as an entire system. This study includes the physical nature of neutrons and their feedback with power generation.  This feedback includes coupling with other physical properties such as temperature, fluids, material dynamics, etc.  This section describes several dynamics examples, including the LRA benchmark and two TREAT experiments.  These examples are of increased complexity from the kinetics examples of \chpt{sect:kin}.  This Chapter also analyzes IQS's performance with these, which is vital for verification of IQS in real-world problems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{LRA Benchmark}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.  It is a super prompt-critical transient.  To have better understanding on the cross sections given later, we present the equations here:
\tcr{ooops, wrong sign on the RHS for neutronics!}
\begin{subequations}
\begin{align}
-\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\div D_2 \grad\phi_2 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
P &= \kappa S_f,
\end{align}
\end{subequations}
where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0(\vec{r})=T(\vec{r}, t=0)$.
The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
\begin{align}
-\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
-\div D_2 \grad\phi_2 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1.
\end{align}
The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.  The initial flux distribution shall be normalized such that the averaged power density
\begin{align}
\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
\end{align}
where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LRA Multiphysics Time Scale Results}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LRA with Time Adaptation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{TREAT Transient-15 Problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Transient 15 is a test case based on the TREAT core. The preliminary purpose of this model was to match an early test of TREAT; due to lack of experimental data and procedures, model validation was impossible \cite{mammoth}. Therefore, ultimate purpose of this simulation in Rattlesnake is to test the model's fidelity with the thermal feedback of TREAT, but it is not meant to exactly match any previous experiments.  Nevertheless, the goal of the following simulations is to test IQS and its time scale based treatment of temperature with a more complex model. The model involves a 159-element "small core" configuration of TREAT, shown in \fig{fig:Tran15_config} \cite{Tran15}. \fig{fig:Tran15_mesh} shows the meshing techniques of the model \cite{mammoth}. To ease computation, the fully homogenized version (without air gaps) was used  for the following simulations. Transient 15 involves an 11-energy group diffusion approximation and is discretized into $355,712$ hexahedral continuous finite elements totaling $4,109,523$ degrees of freedom.  The three-second transient involves a linear ramp decrease in the absorption cross section throughout the control rod region. \fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.   
\tcr{I do not see/understand the air channel meshing}
\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\linewidth]{\FiguresDir/Tran15_config.png}
\caption{Transient-15 159-element small core configuration}
\label{fig:Tran15_config}
\end{figure}

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/Tran15_mesh_fine.png}
\caption{Explicit meshing of air channels}
\label{fig:Tran15_mesh_fine}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/Tran15_mesh_homo.png}
\caption{Fully homogenized fuel elements}
\label{fig:Tran15_mesh_homo}
\end{subfigure}
\caption{Top quarter view of Transient-15 mesh}
\label{fig:Tran15_mesh}
\end{figure}

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.6\linewidth]{\FiguresDir/Tran15_core2.png}
\caption{Transient-15 core power profile at peak power}
\label{fig:Tran15}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transient-15 Temperature Feedback}

@@

\be
c_p = -5.8219\times 10^{-10}T^3 - 4.3694\times 10^{-7}T^2 + 2.8369\times 10^{-3}T -1.009\times 10^{-2} \ (J/g/K)
\label{eq:cp}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transient-15 Multiphysics Time Scale Results}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{\FiguresDir/Tran15_profile.png}
\caption{Transient-15 total power and average temperature profile during transient}
\label{fig:Tran15_profile}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\linewidth]{\FiguresDir/time_constant_tran15.png}
\caption{Dynamical time scale for the Transient-15 example}
\label{fig:tran15tc}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transient-15 with Time Adaptation}

@@

%\section{TREAT M8-CAL Problem}
